Approximation as Differentiation

### 2025-06-18 main/first-order.tex
Some progress, notationally a bit tricky...
Rough pass on why LatGal has a lifting monad.

To interpret the syntactic type constructor $\synVar{L}$, we note that $\LatGal$ has a lifting monad
$\mathcal{L}$ which sends any bounded lattice $X = (X, \meet, \join, \top, \bot)$ to the lattice
$\mathcal{L}(X) = X_\bot$.

The families construction, regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F:
\cat{C} \to \cat{D}$ to functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category
by postcomposing with $F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f,
\partial f)$ to $(f, F \comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where
$(F \comp \partial f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then
$\Fam(F)$ is a (strong) monad on $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to
\cat{C}$ componentwise to families. Thus the lifting monad $\mathcal{L}$ in $\LatGal$ is also a lifting
monad in $\Fam(\LatGal)$.

>>>>

To interpret the syntactic type constructor $\synVar{L}$, consider the strong monad $(\mathcal{L}, \eta, \mu)$
which extends any poset $X$ with a distinguished bottom element $\bot$ and any monotone function $f$ between
posets with a mapping from $\bot$ to $\bot$. $\mathcal{L}$ is also a strong monad on meet-semilattices. On
(bounded) join-semilattices, $\mathcal{L}$ induces a costrong comonad $(\mathcal{L}, \varepsilon, \delta)$,
with the bottom element of $X$ needed to implement the counit $\varepsilon_X: \mathcal{L}(X) \to X$. These two
combine to form a strong lifting monad in $\LatGal$ with unit $\varepsilon_X \dashv \eta_X$ and multiplication
$\delta_X \dashv \mu_X$.

The lifting monad $\mathcal{L}$ in $\LatGal$ is preserved into $\Fam(\LatGal)$. The families construction,
regarded as a covariant 2-functor $\Fam(-): \Cat \to \Cat$, sends functors $F: \cat{C} \to \cat{D}$ to
functors $\Fam(F): \Fam(\cat{C}) \to \Fam(\cat{D})$ that reassign the target category by postcomposing with
$F$, sending objects $(X, \partial X)$ to $(X, F \comp \partial X)$ and morphisms $(f, \partial f)$ to $(f, F
\comp \partial f)$, where $F \comp \partial f$ denotes the natural transformation where $(F \comp \partial
f)(x) = F(\partial f(x))$. If $F$ happens to be a (strong) monad on $\cat{C}$, then $\Fam(F)$ is a (strong)
monad on $\Fam(\cat{C})$, computed by extending the original monad $F: \cat{C} \to \cat{C}$ componentwise to
families.
